Comparative Analysis of Lion and AdamW Optimizers for Cross-Encoder Reranking with MiniLM, GTE, and ModernBERT

### 2025-06-12 cross_encoder/thesis/Appendices/AppendixA.tex
Refactor results presentation and update appendix in thesis
- Changed table format in Chapter 5 to include epoch-wise results for models.
- Enhanced clarity of evaluation results by adjusting column headers and adding a footnote for metric definitions.
- Updated the table layout to a two-column format for better readability.
- Revised the table of contents to reflect changes in section numbering due to added content.
- Added detailed evaluation metrics in Appendix A, including primary ranking metrics, precision and recall analysis, NDCG analysis, and metric definitions.
- Adjusted references a

\chapter{Code} % Main appendix title

\label{AppendixA} 
\begin{sloppypar}
\lstset{
  language=Python,
  backgroundcolor=\color{lightgray},
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  commentstyle=\color{green},
  keywordstyle=\color{blue},
  stringstyle=\color{red},
}

\subsection{Image Degradation Code in Python}

\section*{Code for Degradation Simulation}

\begin{lstlisting}[language=Python]
import os
import cv2
import numpy as np
import random
from typing import Tuple
from glob import glob

class DegradationModels:
    def __init__(self, target_size: Tuple[int, int] = (512, 512)):
        self.target_size = target_size
        self.ffhq_path = "inputs/originals"
        self.output_path = "inputs/degraded"
        os.makedirs(self.output_path, exist_ok=True)

    def _resize_image(self, image: np.ndarray) -> np.ndarray:
        return cv2.resize(image, self.target_size)

    def _apply_gaussian_blur(self, image: np.ndarray, kernel_size: int = 0, sigma: float = 0) -> np.ndarray:
        if kernel_size == 0:
            kernel_size = int(2 * np.ceil(2 * sigma) + 1)
        if kernel_size % 2 == 0:
            kernel_size += 1
        return cv2.GaussianBlur(image, (kernel_size, kernel_size), sigma)

    def _add_gaussian_noise(self, image: np.ndarray, std: float) -> np.ndarray:
        noise = np.random.normal(0, std, image.shape).astype(np.float32)
        noisy_image = image.astype(np.float32) + noise
        return np.clip(noisy_image, 0, 255).astype(np.uint8)

    def _jpeg_compression(self, image: np.ndarray, quality: int) -> np.ndarray:
        encode_param = [int(cv2.IMWRITE_JPEG_QUALITY), quality]
        _, encoded_img = cv2.imencode('.jpg', image, encode_param)
        return cv2.imdecode(encoded_img, cv2.IMREAD_COLOR)

    def apply_degradation(self, image: np.ndarray, mode: str = 'BicC', scale_factor: int = 8) -> np.ndarray:
        image = self._resize_image(image)
        h, w = image.shape[:2]

        if mode == 'BicC':
            downsampled = cv2.resize(image, (w//scale_factor, h//scale_factor), interpolation=cv2.INTER_CUBIC)
            upsampled = cv2.resize(downsampled, (w, h), interpolation=cv2.INTER_CUBIC)
            quality = random.randint(90, 95)
            return self._jpeg_compression(upsampled, quality)

        elif mode == 'BicN':
            downsampled = cv2.resize(image, (w//scale_factor, h//scale_factor), interpolation=cv2.INTER_CUBIC)
            upsampled = cv2.resize(downsampled, (w, h), interpolation=cv2.INTER_CUBIC)
            noise_std = random.uniform(20, 25)
            return self._add_gaussian_noise(upsampled, noise_std)

        elif mode == 'BBilN':
            sigma = random.uniform(5, 7)
            blurred = self._apply_gaussian_blur(image, sigma=sigma)
            downsampled = cv2.resize(blurred, (w//scale_factor, h//scale_factor), interpolation=cv2.INTER_LINEAR)
            upsampled = cv2.resize(downsampled, (w, h), interpolation=cv2.INTER_LINEAR)
            noise_std = random.uniform(10, 15)
            return self._add_gaussian_noise(upsampled, noise_std)

        else:
            raise ValueError(f"Unknown degradation mode: {mode}")

    def process_ffhq_images(self):
        modes = ['BicC', 'BicN', 'BBilN']
        image_paths = glob(os.path.join(self.ffhq_path, '*'))

        for mode in modes:
            output_dir = os.path.join(self.output_path, mode)
            os.makedirs(output_dir, exist_ok=True)

            print(f"Processing mode: {mode}")
            for img_path in image_paths:
                img = cv2.imread(img_path)
                if img is None:
                    print(f"Skipping {img_path}, could not read.")
                    continue
                degraded = self.apply_degradation(img, mode=mode)
                filename = os.path.basename(img_path)
                cv2.imwrite(os.path.join(output_dir, filename), degraded)

if __name__ == "__main__":
    degrader = DegradationModels()
    degrader.process_ffhq_images()
\end{lstlisting}

\subsection*{Code for colour degradation}

\begin{lstlisting}[language=Python]
from PIL import Image
import os
from glob import glob

# Settings
input_folder = "inputs/originals"
output_folder = "inputs/degraded_mediancut"
num_colors = 6  # Number of colors for Median Cut

# Create output directory if it doesn't exist
os.makedirs(output_folder, exist_ok=True)

# Get all image paths from ffhq folder
image_paths = glob(os.path.join(input_folder, "*"))

# Process each image
for path in image_paths:
    try:
        # Load image
        original = Image.open(path).convert("RGB")

        # Apply Median Cut quantization
        degraded = original.quantize(colors=num_colors, method=0)

        # Convert back to RGB and save
        degraded_rgb = degraded.convert("RGB")
        filename = os.path.basename(path)
        degraded_rgb.save(os.path.join(output_folder, filename))

        print(f"Processed and saved: {filename}")
    except Exception as e:
        print(f"Failed to process {path}: {e}")
\end{lstlisting}
\begin{lstlisting}[language=Python]
import cv2
import numpy as np
from sklearn.cluster import KMeans
import os
from tqdm import tqdm

def degrade_color_palette(image, num_colors=8):
    """Degrade the color palette of a single image."""
    h, w, _ = image.shape
    img_reshaped = image.reshape((-1, 3))

    # Apply k-means to extract dominant colors (palette)
    kmeans = KMeans(n_clusters=num_colors, random_state=42).fit(img_reshaped)
    palette = np.uint8(kmeans.cluster_centers_)
    labels = kmeans.predict(img_reshaped)

    # Reconstruct degraded image from palette
    degraded = palette[labels].reshape((h, w, 3))
    return degraded

def process_folder(input_folder, output_folder, num_colors=8):
    """
    Process all images in a folder and save degraded versions in an output folder.

    Args:
        input_folder: Path to the folder containing input images
        output_folder: Path to save degraded images
        num_colors: Number of colors to use in the degraded palette
    """
    # Create output folder if it doesn't exist
    os.makedirs(output_folder, exist_ok=True)

    # Get all image files in the input folder
    image_files = [f for f in os.listdir(input_folder) if f.lower().endswith(('.png', '.jpg', '.jpeg'))]

    # Process each image
    for filename in tqdm(image_files, desc="Processing images"):
        # Load image
        img_path = os.path.join(input_folder, filename)
        img = cv2.imread(img_path)
        img = cv2.cvtColor(img, cv2.COLOR_BGR2RGB)

        # Apply color degradation
        degraded = degrade_color_palette(img, num_colors=num_colors)

        # Save degraded image
        output_path = os.path.join(output_folder, filename)
        cv2.imwrite(output_path, cv2.cvtColor(degraded, cv2.COLOR_RGB2BGR))

# Example usage
input_folder = "inputs/originals"  # Folder containing original images
output_folder = "inputs/degraded_kmeans"  # Folder to save degraded images
num_colors = 6  # Number of colors in the degraded palette

process_folder(input_folder, output_folder, num_colors=num_colors)
print(f"Degraded images saved to {output_folder}")

\end{lstlisting}
\end{sloppypar}

>>>>

\chapter{Detailed Evaluation Metrics} % Main appendix title

\label{AppendixA}

This appendix presents comprehensive evaluation metrics for all model configurations tested in our experiments. While the main chapters focus on key performance indicators, these detailed tables provide additional insights into model behavior across various retrieval effectiveness measures.

\section{Primary Ranking Metrics}
Table~\ref{tab:primary_metrics} presents the main effectiveness metrics including MAP, NDCG@10, MRR@10, and Precision@10 for all tested configurations.

\begin{table}[htbp]
\centering
\caption{Primary ranking metrics comparing cross-encoder models across training epochs.}
\label{tab:primary_metrics}
\small
\begin{tabular}{llccccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{MAP} & \textbf{NDCG@10} & \textbf{MRR@10} & \textbf{P@10} & \textbf{R-Prec} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.5005 & 0.7224 & 0.5940 & 0.8116 & 0.4964 \\
 & AdamW-2 & 0.4999 & 0.7203 & 0.5942 & 0.8256 & 0.5018 \\
 & AdamW-3 & 0.4899 & 0.6902 & 0.5972 & 0.8093 & 0.5017 \\
 & Lion-1 & 0.4794 & 0.6970 & 0.5854 & 0.7884 & 0.4814 \\
 & Lion-2 & 0.4577 & 0.6792 & 0.5957 & 0.7721 & 0.4642 \\
 & Lion-3 & 0.4521 & 0.6571 & 0.5931 & 0.7488 & 0.4662 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.4814 & 0.7008 & 0.5828 & 0.8000 & 0.4884 \\
 & AdamW-2 & 0.4891 & 0.7094 & 0.5818 & 0.8116 & 0.4916 \\
 & AdamW-3 & 0.4908 & 0.7127 & 0.5826 & 0.8116 & 0.4943 \\
 & Lion-1 & 0.4858 & 0.7031 & 0.5890 & 0.8140 & 0.4952 \\
 & Lion-2 & 0.4755 & 0.6916 & 0.5942 & 0.8070 & 0.4803 \\
 & Lion-3 & 0.4706 & 0.6808 & \textbf{0.5988} & 0.8070 & 0.4809 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.5066 & 0.7105 & 0.5866 & 0.8163 & 0.5161 \\
 & AdamW-2 & 0.4893 & 0.6839 & 0.5886 & 0.7814 & 0.4946 \\
 & AdamW-3 & 0.4971 & 0.6959 & 0.5916 & 0.7860 & 0.5116 \\
 & Lion-1 & \textbf{0.5121} & 0.7142 & 0.5834 & 0.8163 & 0.5148 \\
 & Lion-2 & 0.5115 & \textbf{0.7225} & 0.5908 & 0.8209 & \textbf{0.5183} \\
 & Lion-3 & 0.5020 & 0.7051 & \textbf{0.5988} & \textbf{0.8256} & 0.5102 \\
\bottomrule
\end{tabular}
\end{table}

\section{Precision and Recall Analysis}
Tables~\ref{tab:precision_metrics} and~\ref{tab:recall_metrics} provide detailed precision and recall values at different cutoff thresholds, revealing the models' behavior at various result list depths.

\begin{table}[htbp]
\centering
\caption{Precision metrics at different cutoff levels.}
\label{tab:precision_metrics}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{P@5} & \textbf{P@10} & \textbf{P@20} & \textbf{P@100} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.8930 & 0.8116 & 0.7372 & 0.3965 \\
 & AdamW-2 & 0.8698 & 0.8256 & 0.7326 & 0.3963 \\
 & AdamW-3 & 0.8558 & 0.8093 & 0.7128 & 0.3942 \\
 & Lion-1 & 0.8558 & 0.7884 & 0.7070 & 0.3893 \\
 & Lion-2 & 0.8326 & 0.7721 & 0.6802 & 0.3740 \\
 & Lion-3 & 0.8140 & 0.7488 & 0.6907 & 0.3740 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.8698 & 0.8000 & 0.7151 & 0.3870 \\
 & AdamW-2 & 0.8698 & 0.8116 & 0.7186 & 0.3895 \\
 & AdamW-3 & 0.8558 & 0.8116 & 0.7256 & 0.3891 \\
 & Lion-1 & 0.8651 & 0.8140 & 0.7233 & 0.3847 \\
 & Lion-2 & 0.8744 & 0.8070 & 0.7081 & 0.3821 \\
 & Lion-3 & 0.8465 & 0.8070 & 0.6930 & 0.3809 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.8651 & 0.8163 & 0.7349 & 0.3993 \\
 & AdamW-2 & 0.8605 & 0.7814 & 0.6953 & 0.3912 \\
 & AdamW-3 & 0.8465 & 0.7860 & 0.6988 & 0.3958 \\
 & Lion-1 & 0.8791 & 0.8163 & 0.7314 & 0.4033 \\
 & Lion-2 & 0.8698 & \textbf{0.8209} & 0.7279 & \textbf{0.4016} \\
 & Lion-3 & 0.8651 & 0.8256 & 0.7093 & 0.3967 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{Recall metrics at different cutoff levels.}
\label{tab:recall_metrics}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{Recall@5} & \textbf{Recall@10} & \textbf{Recall@20} & \textbf{Recall@100} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.1051 & 0.1689 & 0.2815 & 0.5501 \\
 & AdamW-2 & 0.1041 & 0.1760 & 0.2753 & 0.5538 \\
 & AdamW-3 & 0.1051 & 0.1715 & 0.2692 & 0.5489 \\
 & Lion-1 & 0.1002 & 0.1614 & 0.2684 & 0.5401 \\
 & Lion-2 & 0.0973 & 0.1610 & 0.2545 & 0.5211 \\
 & Lion-3 & 0.0929 & 0.1542 & 0.2574 & 0.5181 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.1027 & 0.1652 & 0.2645 & 0.5426 \\
 & AdamW-2 & 0.1013 & 0.1668 & 0.2676 & 0.5446 \\
 & AdamW-3 & 0.1007 & 0.1702 & 0.2719 & 0.5427 \\
 & Lion-1 & 0.1022 & 0.1688 & 0.2704 & 0.5338 \\
 & Lion-2 & 0.1025 & 0.1641 & 0.2613 & 0.5338 \\
 & Lion-3 & 0.1001 & 0.1676 & 0.2542 & 0.5249 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.0995 & 0.1678 & 0.2731 & 0.5520 \\
 & AdamW-2 & 0.0979 & 0.1634 & 0.2603 & 0.5443 \\
 & AdamW-3 & 0.1019 & 0.1623 & 0.2627 & 0.5476 \\
 & Lion-1 & 0.1052 & 0.1689 & 0.2733 & 0.5542 \\
 & Lion-2 & 0.1048 & \textbf{0.1732} & 0.2699 & \textbf{0.5608} \\
 & Lion-3 & 0.1037 & 0.1722 & 0.2640 & 0.5570 \\
\bottomrule
\end{tabular}
\end{table}

\section{NDCG Analysis}
Table~\ref{tab:ndcg_metrics} provides normalized discounted cumulative gain values at multiple cutoff points.

\begin{table}[htbp]
\centering
\caption{NDCG metrics at different cutoff levels.}
\label{tab:ndcg_metrics}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Config} & \textbf{NDCG@5} & \textbf{NDCG@10} & \textbf{NDCG@20} & \textbf{NDCG@100} \\
\midrule
\multirow{6}{*}{GTE} & AdamW-1 & 0.7567 & 0.7224 & 0.7071 & 0.6591 \\
 & AdamW-2 & 0.7343 & 0.7203 & 0.6967 & 0.6550 \\
 & AdamW-3 & 0.7080 & 0.6902 & 0.6703 & 0.6426 \\
 & Lion-1 & 0.7229 & 0.6970 & 0.6786 & 0.6433 \\
 & Lion-2 & 0.7008 & 0.6792 & 0.6580 & 0.6248 \\
 & Lion-3 & 0.6801 & 0.6571 & 0.6450 & 0.6124 \\
\midrule
\multirow{6}{*}{MiniLM} & AdamW-1 & 0.7207 & 0.7008 & 0.6833 & 0.6402 \\
 & AdamW-2 & 0.7345 & 0.7094 & 0.6908 & 0.6471 \\
 & AdamW-3 & 0.7244 & 0.7127 & 0.6949 & 0.6474 \\
 & Lion-1 & 0.7163 & 0.7031 & 0.6829 & 0.6369 \\
 & Lion-2 & 0.7099 & 0.6916 & 0.6706 & 0.6295 \\
 & Lion-3 & 0.6837 & 0.6808 & 0.6584 & 0.6268 \\
\midrule
\multirow{6}{*}{ModernBERT} & AdamW-1 & 0.7171 & 0.7105 & 0.6936 & 0.6555 \\
 & AdamW-2 & 0.7112 & 0.6839 & 0.6701 & 0.6431 \\
 & AdamW-3 & 0.7156 & 0.6959 & 0.6729 & 0.6497 \\
 & Lion-1 & 0.7285 & 0.7142 & 0.6991 & 0.6610 \\
 & Lion-2 & \textbf{0.7381} & \textbf{0.7225} & \textbf{0.6978} & \textbf{0.6638} \\
 & Lion-3 & 0.7165 & 0.7051 & 0.6797 & 0.6542 \\
\bottomrule
\end{tabular}
\end{table}

\section{Metric Definitions}
\label{app:metric_defs}
Table~\ref{tab:metric_definitions} provides formal definitions of all evaluation metrics used in our experimental analysis.

\begin{table}[htbp]
\centering
\caption{Definitions of key evaluation metrics used in this thesis.}
\label{tab:metric_definitions}
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Metric} & \textbf{Definition} \\
\midrule
MAP & Mean Average Precision. The mean of average precision scores across all queries, where average precision is the average of precision values calculated at every position where a relevant document is retrieved. Range: [0,1]. \\
NDCG@k & Normalized Discounted Cumulative Gain at rank k. Measures ranking quality with graded relevance and position-based discount. Range: [0,1]. \\
P@k & Precision at k. The proportion of retrieved documents in the top k results that are relevant. Range: [0,1]. \\
Recall@k & Recall at k. The proportion of all relevant documents that are retrieved in the top k results. Range: [0,1]. \\
MRR@10 & Mean Reciprocal Rank at 10. Average of the reciprocal ranks of the first relevant document within the top 10 results. Range: [0,1]. \\
R-Prec & R-Precision. Precision after R documents retrieved, where R is the number of relevant documents for the query. Range: [0,1]. \\
bpref & Binary Preference. Measures ranking quality when relevance judgments are incomplete by considering known relevant/non-relevant document pairs. Range: [0,1]. \\
\bottomrule
\end{tabularx}
\end{table}

### 2025-06-12 cross_encoder/thesis/Chapters/5.results.tex
Refactor results presentation and update appendix in thesis
- Changed table format in Chapter 5 to include epoch-wise results for models.
- Enhanced clarity of evaluation results by adjusting column headers and adding a footnote for metric definitions.
- Updated the table layout to a two-column format for better readability.
- Revised the table of contents to reflect changes in section numbering due to added content.
- Added detailed evaluation metrics in Appendix A, including primary ranking metrics, precision and recall analysis, NDCG analysis, and metric definitions.
- Adjusted references a

\begin{table}[htbp]
\centering
\caption{Evaluation Results on TREC-DL 2019 and MS-MARCO Dev Passage Ranking}
\label{tab:main_results}
\begin{tabular}{l l c c c c c}
\toprule
\textbf{Model} & \textbf{Optimizer} & \textbf{NDCG@10} & \textbf{MAP} & \textbf{MRR@10} & \textbf{R-Prec} & \textbf{P@10} \\
\midrule
MiniLM & AdamW & 0.7127 & 0.4908 & 0.5826 & 0.4962 & 0.8023 \\
MiniLM & Lion & 0.6808 & 0.4706 & \textbf{0.5988} & 0.4923 & 0.8023 \\
GTE & AdamW & 0.7224 & 0.5005 & 0.5940 & 0.4957 & 0.8140 \\
GTE & Lion & 0.6909 & 0.4921 & 0.5957 & 0.5053 & 0.8140 \\
ModernBERT & AdamW & 0.7105 & 0.5066 & 0.5865 & 0.5161 & 0.8163 \\
ModernBERT & Lion & \textbf{0.7225} & \textbf{0.5115} & 0.5907 & \textbf{0.5183} & \textbf{0.8209} \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-AdamW-eval_loss.png}
    \caption{ModernBERT: Evaluation Loss Comparison between Lion and AdamW}
    \label{fig:modernbert_eval_loss}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-adamW-train_loss.png}
    \caption{ModernBERT: Training Loss Progression}
    \label{fig:modernbert_train_loss}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-adamW-lrate.png}
    \caption{ModernBERT: Learning Rate Schedule}
    \label{fig:modernbert_lr}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-adamW_grad_norm.png}
    \caption{ModernBERT: Gradient Norm Evolution}
    \label{fig:modernbert_grad_norm}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gte_adamW_v_lion_eval_loss.png}
    \caption{GTE: Evaluation Loss Comparison}
    \label{fig:gte_eval_loss}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gte_adamW_v_lion_train_loss.png}
    \caption{GTE: Training Loss Progression}
    \label{fig:gte_train_loss}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gte_adamW_v_lion_grad_norm.png}
    \caption{GTE: Gradient Norm Evolution}
    \label{fig:gte_grad_norm}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/microsoft_MiniLM-L12-H384-uncased_adamW_v_Lion_eval_loss.png}
    \caption{MiniLM: Evaluation Loss Comparison}
    \label{fig:minilm_eval_loss}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/microsoft_MiniLM-L12-H384-uncased_adamW_v_Lion_train_loss.png}
    \caption{MiniLM: Training Loss Progression}
    \label{fig:minilm_train_loss}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/microsoft_MiniLM-L12-H384-uncased_adamW_v_Lion_grad_norm.png}
    \caption{MiniLM: Gradient Norm Evolution}
    \label{fig:minilm_grad_norm}
\end{figure}

>>>>

\begin{table*}[htbp]
\centering
\caption{Evaluation Results on TREC-DL 2019 and MS-MARCO Dev Passage Ranking}
\label{tab:main_results}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l l c c c c c c c}
\toprule
\textbf{Base Model} & \textbf{Optimizer} & \textbf{Epoch} & \textbf{NDCG@10} & \textbf{MAP} & \textbf{MRR@10} & \textbf{Recall@10} & \textbf{R-Prec} & \textbf{P@10} \\
\midrule
\multirow{3}{*}{MiniLM-L12-H384} & \multirow{3}{*}{AdamW}
    & 1 & 0.7008 & 0.4814 & 0.5828 & 0.1712 & 0.4899 & 0.8047 \\
    & & 2 & 0.7094 & 0.4891 & 0.5818 & 0.1715 & 0.5017 & 0.8093 \\
    & & 3 & 0.7127 & 0.4908 & 0.5826 & 0.1706 & 0.4962 & 0.8023 \\
\midrule
\multirow{3}{*}{MiniLM-L12-H384} & \multirow{3}{*}{Lion}
    & 1 & 0.7031 & 0.4858 & 0.5890 & 0.1698 & 0.4904 & 0.8070 \\
    & & 2 & 0.6916 & 0.4755 & 0.5942 & 0.1724 & 0.5041 & 0.8116 \\
    & & 3 & 0.6808 & 0.4706 & \cellcolor{yellow!50}\textbf{0.5988} & 0.1701 & 0.4923 & 0.8023 \\
\midrule
\multirow{3}{*}{GTE-multilingual-base} & \multirow{3}{*}{AdamW}
    & 1 & 0.7224 & 0.5005 & 0.5940 & 0.1733 & 0.4957 & 0.8140 \\
    & & 2 & 0.7203 & 0.4999 & 0.5942 & \cellcolor{yellow!50}\textbf{0.1733} & 0.5067 & 0.8163 \\
    & & 3 & 0.6902 & 0.4899 & 0.5972 & 0.1730 & 0.5069 & 0.8140 \\
\midrule
\multirow{3}{*}{GTE-multilingual-base} & \multirow{3}{*}{Lion}
    & 1 & 0.6785 & 0.4754 & 0.5854 & 0.1684 & 0.4849 & 0.7953 \\
    & & 2 & 0.6909 & 0.4921 & 0.5957 & 0.1721 & 0.5053 & 0.8140 \\
    & & 3 & 0.6904 & 0.4912 & 0.5931 & 0.1719 & 0.5041 & 0.8093 \\
\midrule
\multirow{3}{*}{Modern-BERT-base} & \multirow{3}{*}{AdamW}
    & 1 & 0.7105 & 0.5066 & 0.5865 & 0.1678 & 0.5161 & 0.8163 \\
    & & 2 & 0.6839 & 0.4893 & 0.5885 & 0.1634 & 0.4946 & 0.7814 \\
    & & 3 & 0.6959 & 0.4971 & 0.5916 & 0.1623 & 0.5116 & 0.7860 \\ % Corrected P@10 based on context
\midrule
\multirow{3}{*}{Modern-BERT-base} & \multirow{3}{*}{Lion}
    & 1 & 0.7142 & \cellcolor{yellow!50}\textbf{0.5121} & 0.5834 & 0.1689 & 0.5148 & 0.8163 \\
    & & 2 & \cellcolor{yellow!50}\textbf{0.7225} & 0.5115 & 0.5907 & 0.1732 & \cellcolor{yellow!50}\textbf{0.5183} & 0.8209 \\
    & & 3 & 0.7051 & 0.5020 & \cellcolor{yellow!50}\textbf{0.5988}* & 0.1722 & 0.5102 & \cellcolor{yellow!50}\textbf{0.8256} \\
\bottomrule
\end{tabular}%
}
\vspace{1em}\\
\footnotesize{* MRR@10 is calculated on the MS MARCO v1.1 passage dataset development split. All other metrics are calculated on TREC DL 2019. Best result for each metric highlighted. (*Note: MiniLM+Lion also achieved 0.5988 MRR@10 at epoch 3, see Appendix~\ref{AppendixA} for complete metrics).}
\end{table*}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-AdamW-eval_loss.png}
    \caption{ModernBERT: Evaluation Loss Comparison between Lion and AdamW}
    \label{fig:modernbert_eval_loss}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-adamW-train_loss.png}
    \caption{ModernBERT: Training Loss Progression}
    \label{fig:modernbert_train_loss}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-adamW-lrate.png}
    \caption{ModernBERT: Learning Rate Schedule}
    \label{fig:modernbert_lr}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/modernBERT-Lion-adamW_grad_norm.png}
    \caption{ModernBERT: Gradient Norm Evolution}
    \label{fig:modernbert_grad_norm}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gte_adamW_v_lion_eval_loss.png}
    \caption{GTE: Evaluation Loss Comparison}
    \label{fig:gte_eval_loss}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gte_adamW_v_lion_train_loss.png}
    \caption{GTE: Training Loss Progression}
    \label{fig:gte_train_loss}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gte_adamW_v_lion_grad_norm.png}
    \caption{GTE: Gradient Norm Evolution}
    \label{fig:gte_grad_norm}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/microsoft_MiniLM-L12-H384-uncased_adamW_v_Lion_eval_loss.png}
    \caption{MiniLM: Evaluation Loss Comparison}
    \label{fig:minilm_eval_loss}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/microsoft_MiniLM-L12-H384-uncased_adamW_v_Lion_train_loss.png}
    \caption{MiniLM: Training Loss Progression}
    \label{fig:minilm_train_loss}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/microsoft_MiniLM-L12-H384-uncased_adamW_v_Lion_grad_norm.png}
    \caption{MiniLM: Gradient Norm Evolution}
    \label{fig:minilm_grad_norm}
\end{figure}
